---
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Peer Grader Guidance
Please review the student expectations for peer review grading and peer review comments.  Overall, we ask that you score with accuracy. When grading your peers, you will not only learn how to improve your future homework submissions but you will also gain deeper understanding of the concepts in the assignments. When assigning scores, consider the responses to the questions given your understanding of the problem and using the solutions as a guide. Moreover, please give partial credit for a concerted effort, but also be thorough. **Add comments to your review, particularly when deducting points, to explain why the student missed the points.** Ensure your comments are specific to questions and the student responses in the assignment.


# Background

You have been contracted as a healthcare consulting company to understand the factors on which the pricing of health insurance depends. 

## Data Description

The data consists of a data frame with 1338 observations on the following 7 variables:

1. price: Response variable ($)
2. age: Quantitative variable
3. sex: Qualitative variable
4. bmi: Quantitative variable
5. children: Quantitative variable
6. smoker: Qualitative variable
7. region: Qualitative variable


## Instructions on reading the data

To read the data in `R`, save the file in your working directory (make sure you have changed the directory if different from the R working directory) and read the data using the `R` function `read.csv()`

```{r}
insurance = read.csv("insurance.csv", head = TRUE)
```


# Question 1: Exploratory Data Analysis [15 points]

a. **4 pts** Create scatterplots of the response, *price*, against three quantitative predictors *age*, *bmi*, and *children*. Describe the general trend (direction and form) of each plot. It should be 3 separate scatter plots.
```{r}

```


b. **4 pts** What is the value of the correlation coefficient for each of the above pair of response and predictor variables? What does it tell you about your comments in part (a)?
```{r}

```


c. **4 pts** Create box plots of the response, *price*, and the three qualitative predictors *sex*, *smoker*, and *region*. Based on these box plots, does there appear to be a relationship between these qualitative predictors and the response?

*Hint*: Use the given code to convert the qualitative predictors to factors.

```{r}
#make categorical variables into factors
insurance$sex<-as.factor(insurance$sex) #makes female the baseline level
insurance$smoker<-as.factor(insurance$smoker) #makes no the baseline level
insurance$region<-as.factor(insurance$region) #makes northeast the baseline level
```


d. **3 pts** Based on the analysis above, does it make sense to run a multiple linear regression with all of the predictors?


# Question 2: Fitting the Multiple Linear Regression Model [12 points]

Build a multiple linear regression model, named *model1*, using the response, *price*, and all 6 predictors, and then answer the questions that follow:

a. **6 pts** Report the coefficient of determination (R-squared) for the model and give a concise interpretation of this value.
```{r}

```


b. **6 pts** Is the model of any use in predicting price? Using $\alpha = 0.05$, provide the following elements of the test of overall regression of the model: null hypothesis $H_0$, alternative hypothesis $H_a$, $F$-statistic or $p$-value, and conclusion.



# Question 3: Model Comparison [14 points]

a. **5 pts** Assuming a marginal relationship between *region* and *price*, perform an ANOVA $F$-test on the mean insurance prices among the different regions. Using an $\alpha$-level of 0.05, can we reject the null hypothesis that the means of the regions are equal? Please interpret.
```{r}

```


b. **5 pts** Now, build a second multiple linear regression model, called *model2*, using *price* as the response variable, and all variables except *region* as the predictors. Conduct a partial $F$-test comparing *model2* with *model1*. What is the partial-F test p-value? Can we reject the null hypothesis that the regression coefficients for *region* variables are zero at an $\alpha$-level of 0.05?
```{r}

```


c. **4 pts** What can you conclude from 3a and 3b? Do they provide the exact same results?


*Note: Please use model1 for all of the following questions.*

# Question 4: Coefficient Interpretation [7 points]

a. **3 pts** Interpret the estimated coefficient of *sexmale* in the context of the problem. *Make sure female is the baseline level for sex. Mention any assumptions you make about other predictors clearly when stating the interpretation.*


b. **4 pts** If the value of the *bmi* in *model1* is increased by 0.01 and the other predictors are kept constant, what change in the response would be expected?


# Question 5: Confidence and Prediction Intervals [12 points]

a. **6 pts** Compute 90% and 95% confidence intervals (CIs) for the parameter associated with *age* for *model1*. What observations can you make about the width of these intervals?
```{r}

```


b. **3 pts** Using *model1*, estimate the average price for all insurance policies with the same characteristics as the first data point in the sample. What is the 95% confidence interval? Provide an interpretation of your results.
```{r}

```


c. **3 pts** Suppose that the *age* value for the first data point is increased to 50, while all other values are kept fixed. Using *model1*, predict the price of an insurance policy with these characteristics. What is the 95% prediction interval? Provide an interpretation of your results.
```{r}

```


